---
title: "Yule–Simpson paradox"
format: 
  html: 
    self-contained: true
    #html-math-method: katex
---

Suppose we have some counts from a binary trial:

```{r, include=FALSE}
counts = rbind(
  treated = c(event = 273, nonevent = 77),
  control = c(event = 289, nonevent = 61)
)
```


```{r}
counts
```


According to each of the following metrics, the risk of the event occurring is reduced for the treatment group relative to the control group.

### Risk difference

```{r, include=FALSE}
riskdiff <- function(counts) {
  treated = counts['treated','event']/sum(counts['treated',])
  control = counts['control','event']/sum(counts['control',])
  treated - control
}
```

$$
\text{risk difference}
  = {\frac{\text{event} \cap \text{treated}}{\text{treated}}}
  - {\frac{\text{event} \cap \text{control}}{\text{control}}}
$$

```{r}
riskdiff(counts)
```

### Risk ratio

$$
\text{risk ratio}
  = {\frac{\text{event} \cap \text{treated}}{\text{treated}}}
  \bigg/ {\frac{\text{event} \cap \text{control}}{\text{control}}}
$$

```{r, include=FALSE}
riskratio <- function(counts) {
  treated = counts['treated','event']/sum(counts['treated',])
  control = counts['control','event']/sum(counts['control',])
  treated/control
}
```

```{r}
riskratio(counts)
```

### Odds ratio

$$
\text{odds ratio}
  = {\frac{\text{treated} \cap \text{event}}{\text{treated} \cap \text{nonevent}}}
  \bigg/ {\frac{\text{control} \cap \text{event}}{\text{control} \cap \text{nonevent}}}
$$

```{r, include=FALSE}
oddsratio <- function(counts) {
  treated = counts['treated','event']/counts['treated','nonevent']
  control = counts['control','event']/counts['control','nonevent']
  treated/control
}
```


```{r}
oddsratio(counts)
```

## Stratification can be misleading

Let's now stratify the population by some variable with values $A$ and $B$:

```{r, include=FALSE}
strata <- list(
  A = rbind(
    treated = c(event = 192, nonevent = 71),
    control = c(event = 55, nonevent = 25)
  ),
  B = rbind(
    treated = c(event = 81, nonevent = 6),
    control = c(event = 234, nonevent = 36)
  )
)
```

```{r}
strata
```

Notice that these stratafied counts partition the original counts...

```{r}
strata$A + strata$B
```

...so you would expect to draw the same overall conclusions about the effectiveness of the treatment, right?